Localization properties, observable character and computational reliability of the linear response in open systems

### 2018-07-06 AnalyticStructure_main.tex
update

\section{Introduction}
Since the days of their foundations, disciplines like Computational Physics or Quantum Chemistry always had to deal with the problem of the 
computational reliability of the results. More specifically, a computational result may be considered as reliable for various reasons.
It might be, for example, in agreement with some previously existing data that the employed computational model is susceptible to reproduce.
The model itself can in this way be classified as ``accurate''; it may then be employed to extract, \textit{in silico},
other quantities that might be then used to \emph{predict} or \emph{interpret} experimental data.
Another concept of reliability is the concept of  ``precision'', which is more focused to the actual results
that a given computational model provides for a specific set of quantities, regardless of their potential agreement with 
experimental data. The aim of precise computational approach is to reduce the \emph{computational} uncertainties
of quantities extracted out of a well-defined model, and provide reference results, to which other computer codes
employing the same model can be compared.

@@

In this paper, we focus our attention to Linear-Response treatments of \emph{molecular} systems, 
in particular referring to LR-TDDFT calculation. We will investigate the impact that 
\emph{open} (i.e. Isolated) Boundary Conditions (BC) that have to be imposed to the system will have on the computational evaluation of LR quantites.
In order to do this, we perform analytic considerations on the equations beyond LR treatment.
Such considerations will enable us to distinguish quantities that \textit{a priori} can be compared among different treatments - and therefore tested for reproducibility - 
from others that would \emph{explicitly} depend of the the employed computational treatment. 
In particular, we will discuss the interplay between reproducible LR quantities and the \emph{locality} of the related objects.
% In order to do this, we perform analytic considerations on the equations beyond LR treatment,
% which will help us identifying how to distinguish -- \textit{a priori} -- quantities that are susceptible to be 
% compared among different treatments, as they are \emph{intrinsically} associated to localized objects,
% from quantities that \emph{explicitly} depend of the the employed computational treatment.

Our considerations will also reveal useful for the understanding of the \emph{analytic structure} of LR quantities like
the linear susceptibility of the system (i.e. the \emph{reducible} polarizability).
Although we refer, as anticipated, to TDDFT models, such considerations might also be transposed to Many-Body Perturbation Theory
approaches, in the context Quasi-Particle equations.

@@

\subsection{Long range behavior of the eigenvalue solution of the Schroedinger equation}
\label{SEopenSystem}
Consider a (one-body) wavefunction $\ket{\psi}$ solution of the equation
$\op H\ket{\psi}=\eps\ket{\psi}$, where we split the Hamiltonian operator into the sum of
a kinetic $\op T$ and a potential $\op V \equiv \op H - \op T$ term.
In view of analysing the impact of the BC, it is instructive to transform 
the Schroedinger Equation in a \emph{scattering} problem, by writing
\be\lb{SEHelmholtz1}
\ket{\psi} = \op G_H(\eps)\ket{\op V\psi}\;,
\ee
where we have introduced the Green function of the Helmholtz operator $\op G_H(\eps) = (\eps-\op T)^{-1}$,
resolvent of the free-particle Hamiltonian $H_s \equiv \op T$.

When describing the Ground State of a molecule, we suppose that the potential operator, even if non-local,
can be effectively restricted on wavefunctions that can be considered as nonzero only on a
bounded domain. In other terms, a molecule can be associated to a ``scatterer'' that has a \emph{finite} 
spatial extension; if a state $\ket{\psi}$ is localized, it will be the same for the ket $\ket{\op V \psi}$, though 
perhaps in a larger region.
For a computational treatment of the problem, this property is fundamental in the 
choice of the numerical treatment; we know that basis functions that are tailored to express asymptotically vanishing 
wavefunctions are better suited for the solution of the problem.

@@

From the above expressions we can see that, \emph{for molecular systems},
the value $\epsilon=0$
represents a \emph{threshold}~\footnote{The threshold value in general should correspond to asymptotic value of the potential at infinity.} 
energy that separates two different classes of solutions.
Negative-energy solutions are \emph{localized}, i.e. they exhibit bound-state behaviour far away from the scatterer.
These are \emph{bound} states and, by means of their normalizability, they are associated to \emph{discrete}, well-defined energies.
The bound states can be of course labelled by quantum numbers that are defined by the properties of the scatterer, namely the
symmetries of the potential $\op V$.
In a computational discretization of the Schroedinger equation, such states can be in principle expressed with arbitrary precision, 
provided that the numerical treatment offers the adequate level of completeness. 
Being discrete and well-identified, the bound-state energies are a property of the molecule and can be associated to \emph{observable}
quantities. The values of their energies can be compared among different treatments and it is in principle possible to
provide reference results on them.

@@

It is easy to see from these concepts that a computational treatment that is tailored for the discretization of bound states
will not be effective for continuum states, due to the different long-range asymptotic behaviour - which in turn is a clear consequence of the 
BC of the problem. In a numerical treatment, discretizing the problem with a \emph{finite}
number of degrees of freedom, we will only have access to \emph{pseudo-} continuum states,
whose energy values and density of states will depend on the representation of the kinetic operator
in the basis set. As a consequence, numerical eigenstates with $\epsilon > 0$ will have a value of their energy that will \emph{always}
depend on the computational setup.

@@

% As a preliminary step for the forthcoming discussion we present a simple argument that allows us to determine the asymptotic character of the eigenvalue solution of the Schroedinger equation (SE) 
% $\op H\ket{\psi}=\eps\ket{\psi}$. The main aims of this analysis, apart from reproducing the well known behavior of the wave functions, is to introduce some tools and setting up notations and 
% terminology that will we widely used in the rest of the paper. 
% 
% The argument proceeds by splitting the Hamiltonian as the sum of the kinetic $\op T$ and potential $\op V$ term. Since we are dealing with molecular-like systems $\op V$ has a compact support when 
% projected in the $\r$-representation. We can formally solve the SE as an inhomogeneous Helmholtz equation by writing
% \be\lb{SEHelmholtz1}
% \ket{\psi} = \op G_H(\eps)\ket{\op V\psi}\;,
% \ee
% where we have introduced the Green function of the Helmholtz operator $\op G_H(\eps) = (\eps-\op T)^{-1}$. We observe that, due to the properties of $\op V$, the \emph{source term} $\ket{\op V\psi}$ 
% of equation \eqref{SEHelmholtz1} is localized in space and, for open boundary conditions in three dimensions, the kernel of $\op G_H(\eps)$ in the coordinate representation reads: 
% \be\lb{HelmholtzKernelDef1}
% \bra{\r} \op G_H(\eps) \ket{\r'} = \frac{1}{4\pi} \begin{cases}
% \frac{e^{-\alpha|\r - \r' |}}{|\r- \r' |} \;,\; & {\textrm for}\; \eps  < 0\; \\ 
% \frac{e^{\ii \alpha |\r-\r' |}}{|\r-\r' |} \;, & {\textrm for} \; \eps \geq 0
% \end{cases} \;,
% \ee
% where $\alpha = \sqrt{2|\eps|}$. 
% This representation of the eigenvalues solution of the SE evidences that the value $\eps=0$ represents the threshold level that determines the asymptotic character of the wave functions. 
% Indeed, for $\eps > 0$,  $\op G_H(\eps)$ behaves as a spherical wave and its convolution with the source term  gives raise to a $\psi(\r)$ delocalized in the whole accessible space. Instead, 
% for values of $\eps < 0$, the kernel contains an exponential damping factor and the associated wave functions exhibit a bound-state behavior in the long range. Finally, we observe that the 
% square-integrability and the continuity of the wave function and its derivatives poses further constrains responsible for the discretization of the energy spectrum for bound-states.  
% 
% The different asymptotic behavior of the wave functions has a deep impact on the possible computational induced error obtained when this quantities are evaluated using a (finite) computational
% setup.

@@

In the contest of the LR the response density satisfies an equation of motion written in the form of a quantum Liouville operator (for example see \cite{baroni2008})
\be\lb{LiouvillianRhopomegaDef1}
\left(\omega - \Liouv\right) \dm'_\Phi(\omega) =  \commutator{\op\Phi(\omega)}{\dmnot} \;,
\ee
where $\op\Phi(\omega)$ represents the perturbing field. 
The inspection of the linear order time evolution of the eigenstates of $\hnot$ evidence that the response density acts as a transition 
operator, linking the occupied and empty subspaces of $\hnot$, and satisfies the \emph{transverse} condition  
\be\lb{RhopTransverseDef1}
\dm'_\Phi = \dm_{\perp} \doteq \dmnot\dm'_\Phi\op Q_0 + \op Q_0\dm'_\Phi\dmnot \;,
\ee
where $\op Q_0=\identity-\dmnot$ is the projector in the subspace orthogonal to the occupied states of $\hnot$. 
Due to this fact we can introduce an explicit representation of the 
response density parametrized as
\be\lb{rhoPrimeFluctuationStateDef1}
\dm'_\Phi(\omega) = \sum_p\left(\ketbra{\psi_p}{f_p^\Phi(-\omega)} + \ketbra{f_p^\Phi(\omega)}{\psi_p}\right) \;.
\ee
Here we have introduced the set of $\omega$-dependent \emph{fluctuation states} 
$\ket{f_p^\Phi(\omega)} = \dm'_\Phi(\omega) \ket{\psi_p}$, 
that have values in the \emph{unoccupied} subspace of $\hnot$. 
An analysis of the Liouville equation \eqref{LiouvillianRhopomegaDef1} for the response density written in this fashion evidences that the equations of motion of fluctuation states are 
written as a modified Sternheimer equation~\cite{mahan1980,giustino2012,giustino2014}
\be\lb{fluctuationStateEqMotion1}
\left[\omega - (\hnot-\eps_p)\right]\ket{f_p^\Phi(\omega)} = \op Q_0(\op\Phi(\omega)+\op V'[\dm'_\Phi](\omega))\ket{\psi_p} \;,
\ee
and we observe that the first-order Hamiltonian contains, 
apart from the perturbing field, a further term $\op V'[\dm'_\Phi]$ due to the density-dependence of $\hnot$.
% The interesting aspect of these equations in the perspective of the present analysis is that they allow us to present a localization argument for the fluctuation states using a procedure  
% analogous to the one depicted in section \ref{SEopenSystem}. So we split the unperturbed Hamiltonian as the sum of the kinetic $\op T$ and potential $\op V$ term and express the solution of 
% \eqref{fluctuationStateEqMotion1} in the form of a self-consistent inhomogeneous Helmholtz equation
It is easy to see that the solution of equation \eqref{fluctuationStateEqMotion1} can be casted in the form of
a set of scattering problems:
%a self-consistent inhomogeneous Helmholtz equation
\be\lb{rhoPrimeFluctuationStateDef2}
\ket{f_p^\Phi(\omega)} = \op G_H(\omega+\epsilon_p)\left(\op V\ket{f_p^\Phi(\omega)} + \ket{s_p^\Phi(\omega)} \right)\;,
\ee
with%:here in this case the source term contains further addend apart from the unperturbed potential and reads:
$%\be
\ket{s_p^\Phi(\omega)} = \op Q_0(\op V'[\dm'_\Phi](\omega) + \op \Phi(\omega))\ket{\psi_p}
$. %\ee

It can be seen that for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level 
that govern the localization properties of the corresponding 
fluctuation state.
As a \emph{stringent} hypothesis, 
let us now assume that the ket $\ket{s_p^\Phi(\omega)}$ is localized, namely its evaluation can be
restricted (if suitable) on a bounded real-space domain.
%The localization character of fluctuation states can be assessed by following the same argument discussed in section \ref{SEopenSystem} for the eigenstates of the SE. In this case 
%the source term depends also from the perturbation added to the system so this analysis applies only to the class 
%of perturbing field for which $\ket{\op\Phi\psi_p}$ is localized, when projected
  %in the $\r$-representation. 
Even when such locality argument holds true, for $\omega > |\eps_p|$, 
the Helmholtz kernel behaves as a spherical wave and  gives raise to a $f_p(\omega,\r)$ which is \emph{delocalized} in the entire real-space domain.
Such delocalization is an intrinsic consequence of the value of $\omega$, and it does not depend on the particular
choice of the operators.
Only for $\omega < |\eps_p|$ the kernel contains an exponential damping factor and $f_p(\omega,\r)$ \emph{may} 
exhibit a bound-state behavior in the long range, once again provided that the above locality arguments on $\ket{s_p^\Phi(\omega)}$ 
-- hence on $\op V'$ and $\Phi$ -- are valid.

Such locality property of the fluctuation state can be directly confirmed for static perturbations at $\omega=0$.
Let us consider a static perturbing operator $\op\Phi_s$ that can be restricted to a finite domain. 
Any local operator, like e.g. a static electric field in the direction $\mathbf F$, $\Phi_s(\r) = -\mathbf F \cdot \r$, would satisfy such property.
We consider the perturbed hamiltonian $\hnot[\rho_{\Phi_s}] + \Phi_s$, where we indicate with $\rho_{\Phi_s}$ the GS density of the
perturbed system. We assume that the perturbation strength ($F$ in the above example) is sufficiently small to be consistent with a LR 
treatment of the problem. 
In this case, the occupied eigenfunction $\ket{\psi_p^{\Phi_s}}$ of the perturbed hamiltonian 
can be extracted with traditional GS techniques, as by the hypothesis of Eq.~\eqref{SEHelmholtz1},
they would exhibit bound-state behavior.\footnote{The fulfilling of Eq.~\eqref{SEHelmholtz1} for energies below threshold
is not \textit{per se} a guarantee of a localized $\ket{\psi}$. With such equation we demonstrate that, \emph{in the case when} a localized
solution exists, this cannot be found for energies above threshold.}
\begin{figure}[t]
\includegraphics[scale=0.68]{CO_statPolvsBox.eps}
\caption{\label{co_alphaStatic}(Color online) Convergence of the statical polarizability (blue continuous line) and of the ground state permanent dipole (red dotted line) 
of the $CO$ molecule with respect to the size of the computational domain. The inset plot evidences the relative percentage convergence using the value associated to the largest 
box as reference.}
\end{figure}

From the linearized Schroedinger Equation of the perturbed problem, it is easy to see that the fluctuation state at $\omega=0$ can be expressed
in terms of the corresponding bound state of the perturbed Hamiltonian, i.e.
$\ket{f_p^{\Phi_s}(0)} \simeq \op Q_0 \ket{\psi_p^{\Phi_s}}$.
For any value of $p$, the locality of the $\ket{f_p^{\Phi_s}(0)}$ therefore directly stems from the
bound-state behavior of $\ket{\psi_p^{\Phi_s}}$. As $\omega=0$ is by hypothesis 
lower than $|\eps_p|$, this is consistent with the above considerations.

% A direct confirmation of the localization character of fluctuation states at $\omega=0$ can be achieved through a ground-state calculation. Indeed, $\ket{f_p(\omega)}$ are nothing but the 
% perturbed KS eigenstates projected out from the domain of $\dmnot$, so for a constant perturbation $\op\Phi_s$ it turns out that $\ket{f_p(0)} = \op Q_0\ket{\psi_p^{\Phi_s}}$. The perturbed KS 
% states $\ket{\psi_p^{\Phi_s}}$ can be computed (beyond LR) by performing a DFT ground-state computation in which the constant perturbing field is added to the external potential and the 
% spatial extension of the fluctuation states can be directly measured. 
In Fig.~\eqref{co_alphaStatic} illustrate this procedure for a $CO$ molecule 
perturbed by a static electric field. The local character of the fluctuation states at $\omega=0$ is 
probed by verifying the convergence of the statical polarizability versus 
the size of the computational domain used to express the fluctuation states. The convergence rate is compared with the 
one of the static dipole of the molecule computed at zero external field. 
The plots evidence that the typical sizes of $\ket{f_p(0)}$ are analogous to the ones of the unperturbed KS orbitals.

@@

This fact has interesting implication in the evaluation of LR quantities.
%
% In the next section we will discuss the implication of the existence of these
% two distinct regimes on the numerical convergence of the observables computed within the LR framework. 
% Before doing this we propose an analogous analysis with the aim to understand the localization 
% properties of the excitation-based description of the LR.
% 
% The localization properties of fluctuation states allows us to assess the reliability of the linear response functional
% computed using the given basis set. 
%
Indeed, writing the response 
density in terms of fluctuation states implies that 
\be\lb{LinearResponseFunctDef2}
\braket{\delta\op O}_\Phi(\omega) = 2\mathrm{Re} \left(\sum_p \bra{\psi_p}\op O\ket{f_p^\Phi(\omega)}\right) \;.
\ee
%so the computation of the linear response functional passes through the evaluation of the scalar products indicated in \eqref{LinearResponseFunctDef2}.
Let us suppose that also the observable of interest $\op O$ exhibit the locality properties of the potentials $ \op V$, $\op \Phi$, $\op V'$.
Once again, it would be enough to consider a local operator for this condition to be met.
We observe that there exists two 
distinct regimes, on the basis of value of $\omega$. 
The first one, which we will refer to as the \emph{below threshold} regime, is realized when
$\omega$ is lower than the ionization potential $|\eps_h|$.
In this case all the fluctuation states are below their threshold level and equation \eqref{LinearResponseFunctDef2} 
is expressed in terms of genuinely localized 
quantities. The quantities $\braket{\psi_p \op O| f_p^\Phi(\omega)}$, 
being scalr products of bound-state-like objects, may thus be evaluated with computational setup that
are similar to the ones usually employed for GS calculations.
This consideration proves why, when a given computer code employs established GS numerical techniques, 
\emph{it is much easier to converge LR quantities for $\omega$ below threshold}.
The convergence in this regime is ``strong'', in the sense that the basis functions employed enable to express 
with arbitrary precision \emph{both} the states $\ket{\op O \psi_p}$ and $\ket{f_p^\Phi(\omega)}$.
This is of course true even when the above states are only implicitly expressed by the employed LR treatment,
 as e.g. in the case of Krylov spaces generated from the states $\ket{\op\Phi(\omega) \psi_p}$ \cite{baroni2006,baroni2008,linlinKPM}.

@@

\subsection{The importance of the continuum sector}
We have presented formal arguments, confirmed by numerical calculations, that show that 
the excitations belonging to the discrete sector of the spectrum have energies
that may be \emph{quantitatively} compared between codes, thanks to their
localized character. 
Such objects are LR quantities, associated to the molecule, whose numerical convergence
might be found by employing techniques which are very similar to 
the one employed for GS calculations. Being expressed by localized excited states,
computational setups with sufficient level of completeness will be such to
express with high \emph{precision} such section of the excitation spectrum.

>>>>

\section{Introduction}
Since the days of their foundations, disciplines like Computational Physics or Quantum Chemistry always had to deal with the problem of the 
computational reliability of the results. More specifically, a computational result may be considered as reliable for various reasons.
It might be, for example, in agreement with some previously existing data that the employed computational model is susceptible to reproduce.
The model itself may in this way be classified as ``accurate''; it can then be employed to extract, \textit{in silico},
other quantities that might be then used to \emph{predict} or \emph{interpret} experimental data.
Another concept of reliability is the concept of  ``precision'', which is more focused to the actual results
that a given computational model provides for a specific set of quantities, regardless of their potential agreement with 
experimental data. The aim of precise computational approach is to reduce the \emph{computational} uncertainties
of quantities extracted out of a well-defined model, and provide reference results, to which other computer codes
employing the same model can be compared.

@@

In this paper, we focus our attention to Linear-Response treatments of \emph{molecular} systems, 
in particular referring to LR-TDDFT calculation. We will investigate the impact that 
\emph{open} (i.e. Isolated) Boundary Conditions (BC) that have to be imposed to the system will have on the computational evaluation of LR quantities.
In order to do this, we perform analytic considerations on the equations beyond LR treatment.
Such considerations will enable us to distinguish quantities that \textit{a priori} can be compared among different treatments - and therefore tested for reproducibility - 
from others that would \emph{explicitly} depend of the the employed computational treatment. 
In particular, we will discuss the interplay between reproducible LR quantities and the \emph{locality} of the related objects.
% In order to do this, we perform analytic considerations on the equations beyond LR treatment,
% which will help us identifying how to distinguish -- \textit{a priori} -- quantities that are susceptible to be 
% compared among different treatments, as they are \emph{intrinsically} associated to localized objects,
% from quantities that \emph{explicitly} depend of the the employed computational treatment.

Our considerations will also reveal useful for the understanding of the \emph{analytic structure} of LR quantities like
the linear susceptibility of the system (i.e. the reducible polarizability).
Although we refer, as anticipated, to TDDFT models, such considerations might also be transposed to Many-Body Perturbation Theory
approaches, in the context Quasi-Particle equations.

@@

\subsection{Long range behavior of the eigenvalue solution of the Schr\"odinger equation}
\label{SEopenSystem}
Consider a (one-body) wavefunction $\ket{\psi}$ solution of the equation
$\op H\ket{\psi}=\eps\ket{\psi}$, where we split the Hamiltonian operator into the sum of
a kinetic $\op T$ and a potential $\op V \equiv \op H - \op T$ term.
In view of analysing the impact of the BC, it is instructive to transform 
the Schr\"odinger equation in a scattering problem, by writing
\be\lb{SEHelmholtz1}
\ket{\psi} = \op G_H(\eps)\ket{\op V\psi}\;,
\ee
where we have introduced the Green function of the Helmholtz operator $\op G_H(\eps) = (\eps-\op T)^{-1}$,
resolvent of the free-particle Hamiltonian $H_s \equiv \op T$.

When describing the Ground State of a molecule, we suppose that the potential operator, even if non-local,
can be effectively restricted on wavefunctions that can be considered as nonzero only on a
bounded domain. In other terms, a molecule can be associated to a ``scatterer'' that has a \emph{finite} 
spatial extension; if a state $\ket{\psi}$ is localized, it will be the same for the ket $\ket{\op V \psi}$, though 
perhaps in a larger region.
For a computational analysis of the problem, this property is fundamental in the choice of the numerical treatment as we know that basis functions 
that are tailored to express asymptotically vanishing wavefunctions are better suited for the solution of the problem.

@@

From the above expressions we can see that, for molecular systems,the value $\epsilon=0$
represents a \emph{threshold}~\footnote{The threshold value in general should correspond to asymptotic value of the potential at infinity.} 
energy that separates two different classes of solutions.
Negative-energy solutions are \emph{localized}, i.e. they exhibit bound-state behavior far away from the scatterer.
These are bound states and, by means of their normalizability, they are associated to \emph{discrete}, well-defined energies.
The bound states can be of course labelled by quantum numbers that are defined by the properties of the scatterer, namely the
symmetries of the potential $\op V$.
In a computational discretization of the Schr\"odinger equation, such states can be in principle expressed with arbitrary precision, 
provided that the numerical treatment offers the adequate level of completeness. 
Being discrete and well-identified, the bound-state energies are a property of the molecule and can be associated to \emph{observable}
quantities. The values of their energies can be compared among different treatments and it is in principle possible to
provide reference results on them.

@@

It is easy to see from these concepts that a computational treatment that is tailored for the discretization of bound states
will not be effective for continuum states, due to the different long-range asymptotic behavior - which in turn is a clear consequence of the 
BC of the problem. In a numerical treatment, discretizing the problem with a \emph{finite}
number of degrees of freedom, we will only have access to a \emph{pseudo}-continuum of states,
whose energy values and density of states will depend on the representation of the kinetic operator
in the basis set. As a consequence, numerical eigenstates with $\epsilon > 0$ will have a value of their energy that will \emph{always}
depend on the computational setup.

@@

% As a preliminary step for the forthcoming discussion we present a simple argument that allows us to determine the asymptotic character of the eigenvalue solution of the Schr\"odinger equation (SE) 
% $\op H\ket{\psi}=\eps\ket{\psi}$. The main aims of this analysis, apart from reproducing the well known behavior of the wave functions, is to introduce some tools and setting up notations and 
% terminology that will we widely used in the rest of the paper. 
% 
% The argument proceeds by splitting the Hamiltonian as the sum of the kinetic $\op T$ and potential $\op V$ term. Since we are dealing with molecular-like systems $\op V$ has a compact support when 
% projected in the $\r$-representation. We can formally solve the SE as an inhomogeneous Helmholtz equation by writing
% \be\lb{SEHelmholtz1}
% \ket{\psi} = \op G_H(\eps)\ket{\op V\psi}\;,
% \ee
% where we have introduced the Green function of the Helmholtz operator $\op G_H(\eps) = (\eps-\op T)^{-1}$. We observe that, due to the properties of $\op V$, the \emph{source term} $\ket{\op V\psi}$ 
% of equation \eqref{SEHelmholtz1} is localized in space and, for open boundary conditions in three dimensions, the kernel of $\op G_H(\eps)$ in the coordinate representation reads: 
% \be\lb{HelmholtzKernelDef1}
% \bra{\r} \op G_H(\eps) \ket{\r'} = \frac{1}{4\pi} \begin{cases}
% \frac{e^{-\alpha|\r - \r' |}}{|\r- \r' |} \;,\; & {\textrm for}\; \eps  < 0\; \\ 
% \frac{e^{\ii \alpha |\r-\r' |}}{|\r-\r' |} \;, & {\textrm for} \; \eps \geq 0
% \end{cases} \;,
% \ee
% where $\alpha = \sqrt{2|\eps|}$. 
% This representation of the eigenvalues solution of the SE evidences that the value $\eps=0$ represents the threshold level that determines the asymptotic character of the wave functions. 
% Indeed, for $\eps > 0$,  $\op G_H(\eps)$ behaves as a spherical wave and its convolution with the source term  gives raise to a $\psi(\r)$ delocalized in the whole accessible space. Instead, 
% for values of $\eps < 0$, the kernel contains an exponential damping factor and the associated wave functions exhibit a bound-state behavior in the long range. Finally, we observe that the 
% square-integrability and the continuity of the wave function and its derivatives poses further constrains responsible for the discretization of the energy spectrum for bound-states.  
% 
% The different asymptotic behavior of the wave functions has a deep impact on the possible computational induced error obtained when this quantities are evaluated using a (finite) computational
% setup.

@@

In the contest of the LR the response density satisfies an equation of motion written in the form of a quantum Liouville operator (for example see \cite{baroni2008})
\be\lb{LiouvillianRhopomegaDef1}
\left(\omega - \Liouv\right) \dm'_\Phi(\omega) =  \commutator{\op\Phi(\omega)}{\dmnot} \;,
\ee
where $\op\Phi(\omega)$ represents the perturbing field. 
The inspection of the linear order time evolution of the eigenstates of $\hnot$ evidence that the response density acts as a transition 
operator, linking the occupied and empty subspaces of $\hnot$, and satisfies the \emph{transverse} condition  
\be\lb{RhopTransverseDef1}
\dm'_\Phi = \dm_{\perp} \doteq \dmnot\dm'_\Phi\op Q_0 + \op Q_0\dm'_\Phi\dmnot \;,
\ee
where $\op Q_0=\identity-\dmnot$ is the projector in the subspace orthogonal to the occupied states of $\hnot$. 
Due to this fact we can introduce an explicit representation of the 
response density parametrized as
\be\lb{rhoPrimeFluctuationStateDef1}
\dm'_\Phi(\omega) = \sum_p\left(\ketbra{\psi_p}{f_p^\Phi(-\omega)} + \ketbra{f_p^\Phi(\omega)}{\psi_p}\right) \;.
\ee
Here we have introduced the set of $\omega$-dependent \emph{fluctuation states} 
$\ket{f_p^\Phi(\omega)} = \dm'_\Phi(\omega) \ket{\psi_p}$, that have values in the unoccupied subspace of $\hnot$. 
An analysis of the Liouville equation \eqref{LiouvillianRhopomegaDef1} for the response density written in this fashion evidences that the equations of motion of fluctuation states are 
written as a modified Sternheimer equation~\cite{mahan1980,giustino2012,giustino2014}
\be\lb{fluctuationStateEqMotion1}
\left[\omega - (\hnot-\eps_p)\right]\ket{f_p^\Phi(\omega)} = \op Q_0(\op\Phi(\omega)+\op V'[\dm'_\Phi](\omega))\ket{\psi_p} \;,
\ee
and we observe that the first-order Hamiltonian contains, 
apart from the perturbing field, a further term $\op V'[\dm'_\Phi]$ due to the density-dependence of $\hnot$.
% The interesting aspect of these equations in the perspective of the present analysis is that they allow us to present a localization argument for the fluctuation states using a procedure  
% analogous to the one depicted in section \ref{SEopenSystem}. So we split the unperturbed Hamiltonian as the sum of the kinetic $\op T$ and potential $\op V$ term and express the solution of 
% \eqref{fluctuationStateEqMotion1} in the form of a self-consistent inhomogeneous Helmholtz equation
It is easy to see that the solution of equation \eqref{fluctuationStateEqMotion1} can be casted in the form of
a set of scattering problems:
%a self-consistent inhomogeneous Helmholtz equation
\be\lb{rhoPrimeFluctuationStateDef2}
\ket{f_p^\Phi(\omega)} = \op G_H(\omega+\epsilon_p)\left(\op V\ket{f_p^\Phi(\omega)} + \ket{s_p^\Phi(\omega)} \right)\;,
\ee
with%:here in this case the source term contains further addend apart from the unperturbed potential and reads:
$%\be
\ket{s_p^\Phi(\omega)} = \op Q_0(\op V'[\dm'_\Phi](\omega) + \op \Phi(\omega))\ket{\psi_p}
$. %\ee

It can be seen that for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level 
that govern the localization properties of the corresponding fluctuation state.
As a \emph{stringent} hypothesis, let us now assume that the ket $\ket{s_p^\Phi(\omega)}$ is localized, namely its evaluation can be
restricted (if suitable) on a bounded real-space domain.
%The localization character of fluctuation states can be assessed by following the same argument discussed in section \ref{SEopenSystem} for the eigenstates of the SE. In this case 
%the source term depends also from the perturbation added to the system so this analysis applies only to the class 
%of perturbing field for which $\ket{\op\Phi\psi_p}$ is localized, when projected
%in the $\r$-representation. 
Even when such locality argument holds true, for $\omega > |\eps_p|$, 
the Helmholtz kernel behaves as a spherical wave and  gives raise to a $\ket{f^\Phi_p(\omega)}$ which is \emph{delocalized} in the entire real-space domain.
Such delocalization is an intrinsic consequence of the value of $\omega$, and does not depend on the particular choice of the operators.
Only for $\omega < |\eps_p|$ the kernel contains an exponential damping factor and $\ket{f^\Phi_p(\omega)}$ exhibits a bound-state behavior in the long range, 
once again provided that the above locality arguments on $\ket{s_p^\Phi(\omega)}$ -- hence on $\op V'$ and $\Phi$ -- are valid.

Such locality property of the fluctuation state can be directly confirmed for static perturbations at $\omega=0$.
Let us consider a static perturbing operator $\op\Phi_s$ that can be restricted to a finite domain. 
Any local operator, like \emph{e.g.} a static electric field $\mathbf F$, with associated scalar potential $\Phi_s(\r) = -\mathbf F \cdot \r$, would satisfy such property.
We consider the perturbed hamiltonian $\hnot[\dm_{\Phi_s}] + \op\Phi_s$, where $\dm_{\Phi_s}$ represents the GS density of the perturbed system, and assume that the 
perturbation strength ($|\mathbf F|$ in the above example) is sufficiently small to be consistent with a LR treatment of the problem. 
In this case, the occupied eigenfunction $\ket{\psi_p^{\Phi_s}}$ of the perturbed hamiltonian can be extracted with traditional GS techniques, as by the hypothesis 
of Eq.~\eqref{SEHelmholtz1}, they exhibit bound-state behavior.\footnote{The fulfilling of Eq.~\eqref{SEHelmholtz1} for energies below threshold
is not \textit{per se} a guarantee of a localized $\ket{\psi}$. With such equation we demonstrate that, \emph{in the case when} a localized
solution exists, this cannot be found for energies above threshold.}
\begin{figure}[t]
\includegraphics[scale=0.68]{CO_statPolvsBox.eps}
\caption{\label{co_alphaStatic}(Color online) Convergence of the statical polarizability (blue continuous line) and of the ground state permanent dipole (red dotted line) 
of the $CO$ molecule with respect to the size of the computational domain. The inset plot evidences the relative percentage convergence using the value associated to the largest 
box as reference.}
\end{figure}

From the linearized Schr\"odinger Equation of the perturbed problem, it is easy to see that the fluctuation state at $\omega=0$ can be expressed
in terms of the corresponding bound state of the perturbed Hamiltonian, i.e.
$\ket{f_p^{\Phi_s}(0)} \simeq \op Q_0 \ket{\psi_p^{\Phi_s}}$.
For any value of $p$, the locality of the $\ket{f_p^{\Phi_s}(0)}$ therefore directly stems from the
bound-state behavior of $\ket{\psi_p^{\Phi_s}}$. As $\omega=0$ is by hypothesis 
lower than $|\eps_p|$, this is consistent with the above considerations.

% A direct confirmation of the localization character of fluctuation states at $\omega=0$ can be achieved through a ground-state calculation. Indeed, $\ket{f_p(\omega)}$ are nothing but the 
% perturbed KS eigenstates projected out from the domain of $\dmnot$, so for a constant perturbation $\op\Phi_s$ it turns out that $\ket{f_p(0)} = \op Q_0\ket{\psi_p^{\Phi_s}}$. The perturbed KS 
% states $\ket{\psi_p^{\Phi_s}}$ can be computed (beyond LR) by performing a DFT ground-state computation in which the constant perturbing field is added to the external potential and the 
% spatial extension of the fluctuation states can be directly measured. 
In Fig.~\eqref{co_alphaStatic} illustrate this procedure for a $CO$ molecule 
perturbed by a static electric field. The local character of the fluctuation states at $\omega=0$ is 
probed by verifying the convergence of the statical polarizability versus 
the size of the computational domain used to express the fluctuation states. The convergence rate is compared with the 
one of the static dipole of the molecule computed at zero external field. 
The plots evidence that the typical sizes of $\ket{f^\Phi_p(0)}$ are analogous to the ones of the unperturbed KS orbitals.

@@

This fact has interesting implication in the evaluation of LR quantities. Indeed, writing the response 
density in terms of fluctuation states implies that 
\be\lb{LinearResponseFunctDef2}
\braket{\delta\op O}_\Phi(\omega) = 2\mathrm{Re} \left(\sum_p \bra{\psi_p}\op O\ket{f_p^\Phi(\omega)}\right) \;.
\ee
%so the computation of the linear response functional passes through the evaluation of the scalar products indicated in \eqref{LinearResponseFunctDef2}.
Let us suppose that also the observable of interest $\op O$ exhibit the locality properties of the potentials $ \op V$, $\op \Phi$, $\op V'$.
Once again, it would be enough to consider a local operator for this condition to be met.
We observe that there exists two 
distinct regimes, on the basis of value of $\omega$. 
The first one, which we will refer to as the \emph{below threshold} regime, is realized when
$\omega$ is lower than the ionization potential $|\eps_h|$.
In this case all the fluctuation states are below their threshold level and equation \eqref{LinearResponseFunctDef2} 
is expressed in terms of genuinely localized 
quantities. The quantities $\braket{\psi_p \op O| f_p^\Phi(\omega)}$, 
being scalar products of bound-state-like objects, may thus be evaluated with computational setup that
are similar to the ones usually employed for GS calculations.
This consideration proves why, when a given computer code employs established GS numerical techniques, 
\emph{it is much easier to converge LR quantities for $\omega$ below threshold}.
The convergence in this regime is ``strong'', in the sense that the basis functions employed enable to express 
with arbitrary precision \emph{both} the states $\ket{\op O \psi_p}$ and $\ket{f_p^\Phi(\omega)}$.
This is of course true even when the above states are only implicitly expressed by the employed LR treatment,
 as e.g. in the case of Krylov spaces generated from the states $\ket{\op\Phi(\omega) \psi_p}$ \cite{baroni2006,baroni2008,linlinKPM}.

@@

%\subsection{The importance of the continuum sector}
\subsection{Physical relevance of the continuum sector}

We have presented formal arguments, confirmed by numerical calculations, that show that 
the excitations belonging to the discrete sector of the spectrum have energies
that can be \emph{quantitatively} compared between codes, thanks to their
localized character. 
Such objects are LR quantities, associated to the molecule, whose numerical convergence
can be found by employing techniques which are very similar to 
the one adopted for GS calculations. Being expressed by localized excited states,
computational setups with sufficient level of completeness are able to
express with high \emph{precision} such section of the excitation spectrum.
